### 2022-01-18 src-latex/parts/30-dataset.tex
Updates from Overleaf
Started to correct and reduce section 3

The directories we have to process are stored in different libraries in Paris\footnote{Citer les bibliothèques sources?}. They have been scanned independently from each other and are available in digital format with various levels of quality. Moreover, they cover a wide period and ave been produced by different publishers and printers. Their contents, indexes, layouts, methods of printing, etc. vary thus from one directory to another (see Figure \ref{fig:directories}).

@@

In this work, we therefore made a selection among the available directories, trying our best to retain copies representative of the diversity of existing directories and of sufficient quality to produce a reliable training data set. The Table \ref{tab:directories} lists the directories selected, their layout, the structure of their entries, and the number of annotated entries according to the indexes considered. For example, the Didot directory of 1854 is organised according to three indexes: by name, by activity, and by street name. The structure of the following entry is characteristic of its index by name: "Batton (D.-A.) *, professeur au Conservatoire de musique et de déclamation, Saint-Georges, 47." This entry begins with a person's name, followed by the initials of their first name in parentheses.The star symbol indicates that this person was awarded the Légion d'Honneur. The following part of this entry represents the person's activity,i.e. their profession, or social status, here "professor at the Conservatory of music and declamation". The street name and number where the person lives or carries out their activity are written at the end of the entry. These pieces of information constituting the structure of the entries are the basic data that we want to extract, deduplicate and structure in order to build a spatio-temporal database of XIX\textsuperscript{th} century Parisian inhabitants and businesses. With the exception of some potentially wordy activity descriptions, most of these pieces of information correspond to named entities. However, as Table \ref{tab:directories} shows, while most entries contain the same types of named entities, the order in which they appear and the way they are written vary from one directory to another. This argues for a supervised, rather than rule-based, NER approach, which requires annotated data.

@@

In order to have a ground truth corpus representative of the diversity of the directories to be processed, we annotated a few pages of each of the major directory series for which we have digitised copies. To that purpose, we used an annotation tool developed for our project and whose graphical user interface (GUI) is presented in Figure \ref{fig:annotator}. This tool loads images of directory pages. Then, it automatically detects each entry in a given page and builds a box around it : these boxes are shown in the left panel of the GUI. Then the text of each entry is extracted using an OCR tool and finally the named entities searched are annotated using a NER tool: the results are displayed respectively in the first and second tabs of the right panel of the GUI. The entry detection, OCR and NER tasks are performed automatically but as they are prone to errors, their results can be corrected manually through the GUI of the annotation tool. In the end, for each page processed, the bounding boxes of the entries that make it up, the exact text contained in each box, and the named entities that are mentioned are stored in a JSON file.

@@

TODO interesting points to detail:
\begin{itemize}
    \item origin
    \item illustration (maybe in first figure)
    \item various numbers (of the original data and of the sample we extracted) : number of entries per pages, variety of directory entries, \dots
    \item statistics about entry lengths (histogram of number of char / tokens par entry)
    \item entities we considered and how we annotated them (how we assessed gold's quality)
    \item statistics on the distribution of entities across directories and their arrangement within entries
    \item OCR data (automated systems and clean text + how we created the "clean" text + whether/how we assessed clean text's quality)
    \item Measure inter-annotator agreement ? 
    \item Explain charset projection
    \item \dots
\end{itemize}

>>>>

The directories to be processed are stored in different libraries in Paris\footnote{\textcolor{green}{Citer les bibliothèques sources?}}. They have been scanned independently and with various levels of quality. Moreover, they cover a wide period and have been produced by different publishers and printers. Therefore, their contents, indexes, layouts, methods of printing, etc. may vary a lot from one directory to another (see Figure \ref{fig:directories}).

@@

In this work, we made a selection among the available directories, trying our best to retain copies representative of the diversity of existing directories and of sufficient quality to produce a reliable training data set. The Table \ref{tab:directories} lists the directories selected, their layout, the structure of their entries, and the number of annotated entries according to the indexes considered. For example, the Didot directory of 1854 is organised according to three indexes: by name, by activity, and by street name. The structure of the following entry is characteristic of its index by name: "Batton (D.-A.) *, professeur au Conservatoire de musique et de déclamation, Saint-Georges, 47." This entry begins with a person's name, followed by the initials of their first name in parentheses.The star symbol indicates that this person was awarded the Légion d'Honneur. The following part of this entry represents the person's activity,i.e. their profession, or social status, here "professor at the Conservatory of music and declamation". The street name and number where the person lives or carries out their activity are written at the end of the entry. These pieces of information constituting the structure of the entries are the basic data that we want to extract, deduplicate and structure in order to build a spatio-temporal database of XIX\textsuperscript{th} century Parisian inhabitants and businesses. With the exception of some potentially wordy activity descriptions, most of these pieces of information correspond to named entities. However, as Table \ref{tab:directories} shows, while most entries contain the same types of named entities, the order in which they appear and the way they are written vary from one directory to another. This argues for a supervised, rather than rule-based, NER approach, which requires annotated data.

@@

To build our ground truth corpus, we annotated a few pages of each of the major directory series for which we have digitised copies, using a tool developed for our project (see Figure \ref{fig:annotator}). First, the image of a given directory page is loaded. Second, entries are automatically detected and boxes are drawn around them as shown in the left panel of the Graphical User Interface (GUI). Then the text of each entry is extracted by an OCR tool and finally named entities are annotated by a NER tool: the results are displayed respectively in the first and second tabs of the right panel of the GUI. The entry detection, OCR and NER tasks are performed automatically. But as they are prone to errors, their results can be corrected manually through the GUI. In the end, the entry bounding boxes, their text, and the named entities are stored in a JSON file.

@@

TODO interesting points to detail:
\begin{itemize}
    \item statistics about entry lengths (histogram of number of char / tokens par entry)
    \item entities we considered and how we annotated them (how we assessed gold's quality)
    \item statistics on the distribution of entities across directories and their arrangement within entries
    \item OCR data (automated systems and clean text + how we created the "clean" text + whether/how we assessed clean text's quality)
    \item Measure inter-annotator agreement ? 
    \item Explain charset projection
    \item \dots
\end{itemize}

### 2022-01-18 src-latex/parts/50-xp-setup.tex
Started to correct and reduce section 3

Next, a NER system extracts the named entities from the text of each entry output by the OCR and the \textbf{NER Q.A.}
is performed. The NER system outputs a text with tags that enclose the entities. To assess the quality of the entity
extraction, we rely on a technique similar as for the OCR evaluation. The predicted text is aligned with the reference
text and the tags are projected in the alignment. Then, the precision, recall and F-measure are computed considering
only the exact matches between entities.

>>>>

Next, a NER system extracts the named entities from the text of each entry output by the OCR and the \textbf{NER Q.A.}
is performed. The NER system outputs a text with tags that enclose the entities. To assess the quality of the entity
extraction, we rely on a technique similar as for the OCR evaluation. The predicted text is aligned with the reference
text and the tags are projected in the alignment. Then, the precision, recall and f-measure (or f-score) are computed considering
only the exact matches between entities. Precision and recall are defined by equation (2); the f-measure in their harmonic mean.

### 2022-01-18 src-latex/parts/60-results.tex
Started to correct and reduce section 3

\begin{table}[h!]
\centering
\caption{NER performances at entry-level for SpaCy NER, CamemBERT and CamemBERT-pretrained when fine-tuned with an increasing number of examples.}
\begin{tabular}{llrrrrrrrr}
       & Number of training examples &  49   &  99   &  199  &  398  &  796  &  1593 &  3186 &  6373 \\
       & \% & 0.8   & 1.6   & 3.1   & 6.2   & 12.5  & 25.0  & 50.0  & 100.0 \\
\midrule\bottomrule
\multirow{3}{*}{\rotatebox{90}{F1 score}} & Camembert &  89.5 &  90.5 &  92.7 &  93.3 &  \textbf{94.1} &  \textbf{94.9} &  \textbf{94.6} &  \textbf{95.1} \\
       & Camembert-pretrained &  \textbf{92.2} &  \textbf{92.9} &  \textbf{93.6} &  \textbf{93.8} &  93.8 &  94.1 &  \textbf{94.6} &  94.4 \\
       & SpaCy NER &  87.0 &  89.0 &  90.3 &  91.9 &  92.1 &  92.8 &  93.2 &  93.5 \\
\cline{1-10}
\multirow{3}{*}{\rotatebox{90}{Precision}} & Camembert &  87.4 &  88.7 &  91.5 &  92.7 &  93.3 &  94.9 &  93.9 &  95.1 \\
       & Camembert-pretrained &  90.8 &  91.8 &  92.9 &  93.0 &  93.0 &  93.4 &  94.1 &  93.9 \\
       & SpaCy NER &  85.6 &  87.7 &  90.0 &  92.0 &  92.4 &  92.8 &  93.1 &  93.7 \\
\cline{1-10}
\multirow{3}{*}{\rotatebox{90}{Recall}} & Camembert &  91.6 &  92.5 &  93.9 &  93.9 &  94.9 &  94.9 &  95.4 &  95.1 \\
       & Camembert-pretrained &  93.6 &  94.0 &  94.4 &  94.6 &  94.6 &  94.8 &  95.0 &  94.9 \\
       & SpaCy NER &  88.6 &  90.4 &  90.7 &  91.7 &  91.9 &  92.8 &  93.3 &  93.4 \\
\end{tabular}
\label{tab:experiment-1-models-performances}
\end{table}

@@

\begin{table}
\caption{CamemBERT and CamemBERT-pretrained NER performances at entry-level on manually corrected and noisy OCRerd entries when fine-tuned on manually corrected entries.}
\begin{tabular}{lrrrrrr}
{} & \multicolumn{3}{c}{Camembert.pretrained} & \multicolumn{3}{c}{Camembert} \\
{} &                  ref & pero-ocr & tesseract &       ref & pero-ocr & tesseract \\
\midrule
Precision &                 96.5 &     93.3 &      90.3 &      96.1 &     92.4 &      89.7 \\
Recall    &                 96.5 &     92.6 &      91.3 &      96.6 &     92.5 &      91.2 \\
F1 score        &                 96.5 &     92.9 &      90.8 &      96.4 &     92.5 &      90.5 \\
\bottomrule
\end{tabular}
\end{table}

>>>>

\begin{table}[h!]
\centering
\caption{NER performances at entry-level for SpaCy NER, CamemBERT and CamemBERT-pretrained when fine-tuned with an increasing number of examples.}
\begin{tabular}{llrrrrrrrr}
       & Training examples &  49   &  99   &  199  &  398  &  796  &  1593 &  3186 &  6373 \\
       & \% & 0.8   & 1.6   & 3.1   & 6.2   & 12.5  & 25.0  & 50.0  & 100.0 \\
\midrule\bottomrule
\multirow{3}{*}{\rotatebox{90}{F1 score}} & Camembert &  89.5 &  90.5 &  92.7 &  93.3 &  \textbf{94.1} &  \textbf{94.9} &  \textbf{94.6} &  \textbf{95.1} \\
       & Camembert-pretrained &  \textbf{92.2} &  \textbf{92.9} &  \textbf{93.6} &  \textbf{93.8} &  93.8 &  94.1 &  \textbf{94.6} &  94.4 \\
       & SpaCy NER &  87.0 &  89.0 &  90.3 &  91.9 &  92.1 &  92.8 &  93.2 &  93.5 \\
\cline{1-10}
\multirow{3}{*}{\rotatebox{90}{Precision}} & Camembert &  87.4 &  88.7 &  91.5 &  92.7 &  93.3 &  94.9 &  93.9 &  95.1 \\
       & Camembert-pretrained &  90.8 &  91.8 &  92.9 &  93.0 &  93.0 &  93.4 &  94.1 &  93.9 \\
       & SpaCy NER &  85.6 &  87.7 &  90.0 &  92.0 &  92.4 &  92.8 &  93.1 &  93.7 \\
\cline{1-10}
\multirow{3}{*}{\rotatebox{90}{Recall}} & Camembert &  91.6 &  92.5 &  93.9 &  93.9 &  94.9 &  94.9 &  95.4 &  95.1 \\
       & Camembert-pretrained &  93.6 &  94.0 &  94.4 &  94.6 &  94.6 &  94.8 &  95.0 &  94.9 \\
       & SpaCy NER &  88.6 &  90.4 &  90.7 &  91.7 &  91.9 &  92.8 &  93.3 &  93.4 \\
\end{tabular}
\label{tab:experiment-1-models-performances}
\end{table}

@@

\begin{table}
\caption{CamemBERT and CamemBERT-pretrained NER performances at entry-level on manually corrected and noisy OCRerd entries when fine-tuned on manually corrected entries.}
\begin{tabular}{lrrrrrrrrr}
Model & \multicolumn{3}{c}{Camembert.pretrained} & \multicolumn{3}{c}{Camembert} & \multicolumn{3}{c}{Camembert.pretrained} \\
Train set & \multicolumn{3}{c}{ref} & \multicolumn{3}{c}{ref} & \multicolumn{3}{c}{pero-ocr} \\
Test set &                  ref & pero-ocr & tesseract &       ref & pero-ocr & tesseract &                  ref & pero-ocr & tesseract \\
\midrule
Precision &                 96.5 &     93.3 &      90.3 &      96.1 &     92.4 &      89.7 &                 94.3 &     93.6 &      89.9 \\
Recall    &                 96.5 &     92.6 &      91.3 &      96.6 &     92.5 &      91.2 &                 96.1 &     94.5 &      92.5 \\
F1 score        &                 96.5 &     92.9 &      90.8 &      96.4 &     92.5 &      90.5 &                 95.2 &     94.1 &      91.2 \\
\end{tabular}
\end{table}
